Structural Divergence Analysis

### 2017-02-20 analysis-structure-core.Rmd
cambios 20/2/2017

```{r load_packages}
# load packages
library(ggplot2)
library(reshape2)
library(plyr)
library(quantreg)
library(bio3d)

@@

```{r set_my_colors}
# set my colors
myColors = c("blue", "darkgreen", "red", "red") 
names(myColors) <- c("exp", "mut", "strong", "mut+sel")
```

```{r set_run_variables}
# set significance factors
f95 = 1.96
f99 = 2.6
f999 = 3.3
```

```{r set_file_names}
# set input filenames
## data
reference.fname = file.path(data.dir, paste(family, "_ref.txt", sep = ""))
protein_list.fname = file.path(data.dir, paste(family, "_list.txt", sep = ""))
site.info.fname = file.path(data.dir, paste(p.ref, "_m.da.ca.csv", sep = ""))
min.da.CM.CA.fname = file.path(data.dir, paste(p.ref, "_min.da.CM.ca.csv", sep = ""))

@@

```{r set_reference_protein}
# which is the reference protein?
reference = as.character(read.table(reference.fname, header = F)[1, 1])
if (reference != p.ref) {
  print("Warning: different references")
}
```

@@

```{r set_experimental_protein_names, comment = ""}
# experimental protein names
protein.exp = read.table(protein_list.fname, header = F, stringsAsFactors = F)[, 1]
protein.exp = protein.exp[protein.exp != reference]
print(protein.exp)
```

```{r read_the_core, comment = ""}
# indexes.core
indexes.core = as.numeric(read.csv(file.path("DATA", paste("indexes.core", "_", family, ".csv", sep = "")))[1, -1])
```

```{r set_site_info}
# read site information of the reference protein
site.info = read.csv(site.info.fname, header = T, sep = ",")
m.min.da.CM.CA = read.csv(min.da.CM.CA.fname)

@@

# read reference pdb and calculate resno, AA, and CN
reference.pdb.fname = file.path(data.dir,paste(reference, ".pdb", sep = ""))
ref.pdb = read.pdb(reference.pdb.fname)
inds = atom.select(ref.pdb, elety = "CA")
ref.pdb.ca = ref.pdb$atom[inds$atom,]
cmap.ca = cmap(ref.pdb$xyz[inds$xyz], dcut = R0, scut = 0, mask.lower = F)
CN = rowSums(cmap.ca, na.rm = T) - 1 # diag(cmap.ca) returned by cmap is 1!
AA = ref.pdb.ca$resid
resno = ref.pdb.ca$resno
pdb.info = data.frame(resno, AA, CN)
names(pdb.info) = c("resno", "AA", paste("CN", as.character(R0), sep = ""))

@@

```{r read_data}
# read data
## exp
dri2.exp = read.csv(dri2.exp.fname, header = T)[, indexes.core]
rownames(dri2.exp) = protein.exp

@@

```{r prepare_site_data}
# prepare the data
## exp
tdri2.exp = as.data.frame(t(as.matrix(dri2.exp)))
tdri2.exp = cbind(site.info, tdri2.exp)
siteComparison.exp = melt(tdri2.exp, id.vars = names(site.info), variable.name = "protein", value.name = "dri2")

@@

```{r mean_se_z.RSD_vs_site, fig.cap = "Black lines are the active sites and grey lines are their contacts"}
# comparison of z.RSD profiles with functional sites
## prepare dat
dat = siteComparison.all.datasets
dat$dataset = revalue(dat$dataset, c("strong" = "mut+sel"))

@@

```{r-comparison-mean-z.RSD-vs-site, fig.cap = "Black lines are the active sites and grey lines are their contacts"}

@@

```{r-comparison-da}

# comparison of means.z.scores with the minimum distance to the functional sites
plot(y = d.exp$mean.z.RSD - d.mut$mean.z.RSD, x = m.min.da.CM.CA$min.da.ca[indexes.core], xlab = "minimal distance to active site (between CAs)", ylab = "exp.mean.z.RSD - mut.mean.z.RSD")
plot(y = d.exp$mean.z.RSD - d.mut$mean.z.RSD, x = m.min.da.CM.CA$min.da.CM[indexes.core], xlab = "minimal distance to active site (between CMs)", ylab = "exp.mean.z.RSD - mut.mean.z.RSD")

## calculate correlations between profiles
r.min.da.CA = cor((d.exp$mean.z.RSD - d.mut$mean.z.RSD), m.min.da.CM.CA$min.da.ca[indexes.core])
r.min.da.CM = cor((d.exp$mean.z.RSD - d.mut$mean.z.RSD), m.min.da.CM.CA$min.da.CM[indexes.core])

>>>>

```{r load-packages}
# load packages
library(ggplot2)
library(reshape2)
library(plyr)
library(quantreg)
library(bio3d)

@@

```{r set-my-colors}
# set my colors
myColors = c("blue", "darkgreen", "red", "red") 
names(myColors) <- c("exp", "mut", "strong", "mut+sel")
```

```{r set-run-variables}
# set significance factors
f95 = 1.96
f99 = 2.6
f999 = 3.3
```

```{r set-file-names}
# set input filenames
## data
reference.fname = file.path(data.dir, paste(family, "_ref.txt", sep = ""))
protein_list.fname = file.path(data.dir, paste(family, "_list.txt", sep = ""))
site.info.fname = file.path(data.dir, paste(p.ref, "_m.da.ca.csv", sep = ""))
min.da.CM.CA.fname = file.path(data.dir, paste(p.ref, "_min.da.CM.ca.csv", sep = ""))

@@

```{r set-reference-protein}
# which is the reference protein?
reference = as.character(read.table(reference.fname, header = F)[1, 1])
if (reference != p.ref) {
  print("Warning: different references")
}
```

@@

```{r set-experimental-protein-names, comment = ""}
# experimental protein names
protein.exp = read.table(protein_list.fname, header = F, stringsAsFactors = F)[, 1]
protein.exp = protein.exp[protein.exp != reference]
print(protein.exp)
```

```{r read-the-core, comment = ""}
# indexes.core
indexes.core = as.numeric(read.csv(file.path("DATA", paste("indexes.core", "_", family, ".csv", sep = "")))[1, -1])
```

```{r set-site-info}
# read site information of the reference protein
site.info = read.csv(site.info.fname, header = T, sep = ",")
m.min.da.CM.CA = read.csv(min.da.CM.CA.fname)

@@

# read reference pdb and calculate resno, AA, and CN
reference.pdb.fname = file.path(data.dir,paste(reference, ".pdb", sep = ""))
ref.pdb = read.pdb(reference.pdb.fname)
inds = atom.select(ref.pdb, elety = "CA")
ref.pdb.ca = ref.pdb$atom[inds$atom,]
cmap.ca = cmap(ref.pdb$xyz[inds$xyz], dcut = R0, scut = 0, mask.lower = F)
CN = rowSums(cmap.ca, na.rm = T) - 1 # diag(cmap.ca) returned by cmap is 1!
n.sites = length(CN)
AA = ref.pdb.ca$resid
resno = ref.pdb.ca$resno
pdb.info = data.frame(resno, AA, CN)
names(pdb.info) = c("resno", "AA", paste("CN", as.character(R0), sep = ""))

@@

```{r read-data}
# read data
## exp
dri2.exp = read.csv(dri2.exp.fname, header = T)[, indexes.core]
rownames(dri2.exp) = protein.exp

@@

```{r prepare-site-data}
# prepare the data
## exp
tdri2.exp = as.data.frame(t(as.matrix(dri2.exp)))
tdri2.exp = cbind(site.info, tdri2.exp)
siteComparison.exp = melt(tdri2.exp, id.vars = names(site.info), variable.name = "protein", value.name = "dri2")

@@

```{r mean-se-z.RSD-vs-site, fig.cap = "Black lines are the active sites and grey lines are their contacts"}
# comparison of z.RSD profiles with functional sites
## prepare dat
dat = siteComparison.all.datasets
dat$dataset = revalue(dat$dataset, c("strong" = "mut+sel"))

@@

```{r comparison-mean-z.RSD-vs-site, fig.cap = "Black lines are the active sites and grey lines are their contacts"}

@@

```{r comparison-min.da}

# comparison of means.z.scores with the minimum distance to the functional sites
plot(y = d.exp$mean.z.RSD - d.mut$mean.z.RSD, x = m.min.da.CM.CA$min.da.CA[indexes.core], xlab = "minimal distance to active site (between CAs)", ylab = "exp.mean.z.RSD - mut.mean.z.RSD")
plot(y = d.exp$mean.z.RSD - d.mut$mean.z.RSD, x = m.min.da.CM.CA$min.da.CM[indexes.core], xlab = "minimal distance to active site (between CMs)", ylab = "exp.mean.z.RSD - mut.mean.z.RSD")

## calculate correlations between profiles
r.min.da.CA = cor((d.exp$mean.z.RSD - d.mut$mean.z.RSD), m.min.da.CM.CA$min.da.CA[indexes.core])
r.min.da.CM = cor((d.exp$mean.z.RSD - d.mut$mean.z.RSD), m.min.da.CM.CA$min.da.CM[indexes.core])

### 2017-02-20 analysis-structure.Rmd
cambios 20/2/2017

```{r set_my_colors}
# set my colors
myColors = c("blue", "darkgreen", "red", "red") 
names(myColors) <- c("exp", "mut", "strong", "mut+sel")
```

```{r set_run_variables}
# set significance factors
f95 = 1.96
f99 = 2.6
f999 = 3.3
```

```{r set_file_names}
# set input filenames
## data
reference.fname = file.path(data.dir, paste(family, "_ref.txt", sep = ""))
protein_list.fname = file.path(data.dir, paste(family, "_list.txt", sep = ""))
site.info.fname = file.path(data.dir, paste(p.ref, "_m.da.ca.csv", sep = ""))
min.da.CM.CA.fname = file.path(data.dir, paste(p.ref, "_min.da.CM.ca.csv", sep = ""))
functional.sites.fname = file.path("DATA", paste(p.ref, "_functionalSites.csv", sep = ""))

## profiles to compare
dri2.exp.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.exp.norm.dr.squarei.csv", sep = ""))
dri2.strong.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_strong.sel_K.analysis_Keff_out_m.theo.norm.dr.squarei.csv", sep = ""))
dri2.mut.fname = file.path(data.dir,paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.theo.norm.dr.squarei.csv", sep = ""))
```

```{r set_reference_protein}
# which is the reference protein?
reference = as.character(read.table(reference.fname, header = F)[1, 1])
if (reference != p.ref) {
  print("Warning: different references")
}
```

```{r read_functional_sites}
# which are the functional sites?
functional.sites = read.csv(functional.sites.fname, sep = ";")
active.site.index = functional.sites$index
description = functional.sites$description
function.info = data.frame(active.site.index, description)
```

We studied the `r family` family. 
We used the ENM `r enm` with R0 = `r R0` to generate mutants.
We used as reference for the ENM the protein with pdb code `r reference`. 
The active.sites sites of the reference are:
```{r}
print(function.info)
```

@@

```{r set_experimental_protein_names, comment = ""}
# experimental protein names
protein.exp = read.table(protein_list.fname, header = F, stringsAsFactors = F)[, 1]
protein.exp = protein.exp[protein.exp != reference]
print(protein.exp)
```

```{r set_site_info}
# read site information of the reference protein
site.info = read.csv(site.info.fname, header = T, sep = ",")
m.min.da.CM.CA = read.csv(min.da.CM.CA.fname)

@@

```{r read_data}
# read data
## exp
dri2.exp = read.csv(dri2.exp.fname, header = T)
site = site.info$site
dri2.exp = dri2.exp[1:length(protein.exp), site]
rownames(dri2.exp) = protein.exp

@@

```{r-mean_se_z.RSD_vs_site, fig.cap = "Black lines are the active sites and grey lines are their contacts"}
# comparison of z.RSD profiles with functional sites
## prepare dat
dat = siteComparison.all.datasets
dat$dataset = revalue(dat$dataset, c("strong" = "mut+sel"))

@@

### Correlation between site profiles

```{r-comparison-mean-z.RSD-vs-site}

@@

```{r-comparisons-between-plots}

@@

```{r-comparison-da}

# comparison of means.z.scores with the minimum distance to the functional sites
plot(y = d.exp$mean.z.RSD - d.mut$mean.z.RSD, x = m.min.da.CM.CA$min.da.ca, xlab = "minimal distance to active site (between CAs)", ylab = "exp.mean.z.RSD - mut.mean.z.RSD")
plot(y = d.exp$mean.z.RSD - d.mut$mean.z.RSD, x = m.min.da.CM.CA$min.da.CM, xlab = "minimal distance to active site (between CMs)", ylab = "exp.mean.z.RSD - mut.mean.z.RSD")

## calculate correlations between profiles
r.min.da.CA = cor((d.exp$mean.z.RSD - d.mut$mean.z.RSD), m.min.da.CM.CA$min.da.ca)
r.min.da.CM = cor((d.exp$mean.z.RSD - d.mut$mean.z.RSD), m.min.da.CM.CA$min.da.CM)

@@

```{r-write-pdbs} 
# write pdb files with mean.z.RSDi as b-factor for building images

@@

>>>>

```{r set-my-colors}
# set my colors
myColors = c("blue", "darkgreen", "red", "red") 
names(myColors) <- c("exp", "mut", "strong", "mut+sel")
```

```{r set-run-variables}
# set significance factors
f95 = 1.96
f99 = 2.6
f999 = 3.3
```

```{r set-file-names}
# set input filenames
## data
reference.fname = file.path(data.dir, paste(family, "_ref.txt", sep = ""))
protein_list.fname = file.path(data.dir, paste(family, "_list.txt", sep = ""))
site.info.fname = file.path(data.dir, paste(p.ref, "_m.da.ca.csv", sep = ""))
min.da.CM.CA.fname = file.path(data.dir, paste(p.ref, "_min.da.CM.ca.csv", sep = ""))
sum.inv.da.CM.CA.fname = file.path(data.dir, paste(p.ref, "_sum.inv.da.CM.ca.csv", sep = ""))
functional.sites.fname = file.path("DATA", paste(p.ref, "_functionalSites.csv", sep = ""))

## profiles to compare
dri2.exp.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.exp.norm.dr.squarei.csv", sep = ""))
dri2.strong.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_strong.sel_K.analysis_Keff_out_m.theo.norm.dr.squarei.csv", sep = ""))
dri2.mut.fname = file.path(data.dir,paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.theo.norm.dr.squarei.csv", sep = ""))

## consurf
score.fname = file.path(data.dir, paste(p.ref, "_consurf.csv", sep = ""))
sim.data.no.sel.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_no.sel_out_df.data.csv", sep = ""))
sim.data.strong.sel.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_strong.sel_out_df.data.csv", sep = ""))

```

```{r set-reference-protein}
# which is the reference protein?
reference = as.character(read.table(reference.fname, header = F)[1, 1])
if (reference != p.ref) {
  print("Warning: different references")
}
```

```{r read-functional-sites}
# which are the functional sites?
functional.sites = read.csv(functional.sites.fname, sep = ";")
active.site.index = functional.sites$index
description = functional.sites$description
function.info = data.frame(active.site.index, description)
```

In this case we studied the `r family` family. 
We used the ENM `r enm` with R0 = `r R0` to generate mutants.
We used as reference for the ENM the protein with pdb code `r reference`. 
The active/functional sites of the reference are:

```{r print-info}
print(function.info)
```

@@

```{r set-experimental-protein-names}
# experimental protein names
protein.exp = read.table(protein_list.fname, header = F, stringsAsFactors = F)[, 1]
protein.exp = protein.exp[protein.exp != reference]
print(protein.exp)
```

```{r set-site-info}
# read site information of the reference protein
site.info = read.csv(site.info.fname, header = T, sep = ",")
m.min.da.CM.CA = read.csv(min.da.CM.CA.fname)
m.sum.inv.da.CM.CA = read.csv(sum.inv.da.CM.CA.fname)

@@

```{r read-data}
# read data
## exp
dri2.exp = read.csv(dri2.exp.fname, header = T)
site = site.info$site
dri2.exp = dri2.exp[1:length(protein.exp), site]
rownames(dri2.exp) = protein.exp

@@

```{r-mean-se-z.RSD-vs-site, fig.cap = "Black lines are the active sites and grey lines are their contacts"}
# comparison of z.RSD profiles with functional sites
## prepare dat
dat = siteComparison.all.datasets
dat$dataset = revalue(dat$dataset, c("strong" = "mut+sel"))

@@

### Correlation between profiles

```{r comparison-mean-z.RSD-vs-site}

@@

### Comparison between profiles

```{r comparisons-between-plots}

@@

```{r comparison-min.da}

# comparison of means.z.scores with the minimum distance to the functional sites
plot(y = d.exp$mean.z.RSD - d.mut$mean.z.RSD, x = m.min.da.CM.CA$min.da.CA, xlab = "minimal distance to active site (between CAs)", ylab = "exp.mean.z.RSD - mut.mean.z.RSD")
plot(y = d.exp$mean.z.RSD - d.mut$mean.z.RSD, x = m.min.da.CM.CA$min.da.CM, xlab = "minimal distance to active site (between CMs)", ylab = "exp.mean.z.RSD - mut.mean.z.RSD")

## calculate correlations between profiles
r.min.da.CA = cor((d.exp$mean.z.RSD - d.mut$mean.z.RSD), m.min.da.CM.CA$min.da.CA)
r.min.da.CM = cor((d.exp$mean.z.RSD - d.mut$mean.z.RSD), m.min.da.CM.CA$min.da.CM)

@@

```{r comparison-sum.inv.da}

# comparison of means.z.scores with the sum.inv distance to the functional sites
plot(y = d.exp$mean.z.RSD - d.mut$mean.z.RSD, x = m.sum.inv.da.CM.CA$sum.inv.da.CA, xlab = "sum.inv distance to active site (between CAs)", ylab = "exp.mean.z.RSD - mut.mean.z.RSD")
plot(y = d.exp$mean.z.RSD - d.mut$mean.z.RSD, x = m.sum.inv.da.CM.CA$sum.inv.da.CM, xlab = "sum.inv distance to active site (between CMs)", ylab = "exp.mean.z.RSD - mut.mean.z.RSD")

## calculate correlations between profiles
r.sum.inv.da.CA = cor((d.exp$mean.z.RSD - d.mut$mean.z.RSD), m.sum.inv.da.CM.CA$sum.inv.da.CA)
r.sum.inv.da.CM = cor((d.exp$mean.z.RSD - d.mut$mean.z.RSD), m.sum.inv.da.CM.CA$sum.inv.da.CM)

cor.df = data.frame("comparison" = c("(exp - mut) vs sum.inv.da.CA", "(exp - mut) vs sum.inv.da.CM"), "R" = c(r.sum.inv.da.CA, r.sum.inv.da.CM))

knitr::kable(cor.df, digits = 2, caption = "Correlation between differences in <Z-scores> profiles and the sum.inv distance to the active site")
```

```{r correlation-da-CN} 
## calculate correlations between profiles
r.min.da.CA.CN = cor(CN, m.min.da.CM.CA$min.da.CA)
r.min.da.CM.CN = cor(CN, m.min.da.CM.CA$min.da.CM)

cor.df = data.frame("comparison" = c("CN vs min.da.CA", "CN vs min.da.CM"), "R" = c(r.min.da.CA.CN, r.min.da.CM.CN))

knitr::kable(cor.df, digits = 2, caption = "Correlation between CN and the minimal distance to the active site")

## calculate correlations between profiles
r.sum.inv.da.CA.CN = cor(CN, m.sum.inv.da.CM.CA$sum.inv.da.CA)
r.sum.inv.da.CM.CN = cor(CN, m.sum.inv.da.CM.CA$sum.inv.da.CM)

cor.df = data.frame("comparison" = c("CN vs sum.inv.da.CA", "CN vs sum.inv.da.CM"), "R" = c(r.sum.inv.da.CA.CN, r.sum.inv.da.CM.CN))

knitr::kable(cor.df, digits = 2, caption = "Correlation between CN and the sum.inv distance to the active site")
```

```{r write-pdbs} 
# write pdb files with mean.z.RSDi as b-factor for building images

@@

### sequence comparison
```{r}
consurf = read.csv(score.fname, sep = ";")
score = consurf$SCORE

sim.data.no.sel = read.csv(sim.data.no.sel.fname)
mut.site.no.sel = sim.data.no.sel$site
p.accept.no.sel = sim.data.no.sel$p.accept

sim.data.strong.sel = read.csv(sim.data.strong.sel.fname)
mut.site.strong.sel = sim.data.strong.sel$site
p.accept.strong.sel = sim.data.strong.sel$p.accept

mean.p.accept.no.sel = c()
mean.p.accept.strong.sel = c()

if (data.dir == "OUT/out_subset_CM_ANM") {
  mut.site.no.sel = mut.site.no.sel - n.sites
  mut.site.strong.sel = mut.site.strong.sel - n.sites
  }

for (i in (1:n.sites)) {
  p.accept.no.sel.i = p.accept.no.sel[which(mut.site.no.sel == i)]
  mean.p.accept.no.sel.i = mean(p.accept.no.sel.i)
  mean.p.accept.no.sel[i] = mean.p.accept.no.sel.i

  p.accept.strong.sel.i = p.accept.strong.sel[which(mut.site.strong.sel == i)]
  mean.p.accept.strong.sel.i = mean(p.accept.strong.sel.i)
  mean.p.accept.strong.sel[i] = mean.p.accept.strong.sel.i
}

par(pty = "s")

plot(x = mean.p.accept.no.sel, y = CN)
plot(x = mean.p.accept.no.sel, y = score)
plot(x = score, y = CN)

r.mean.p.accept.no.sel.CN = cor(mean.p.accept.no.sel, CN)
r.mean.p.accept.no.sel.score = cor(mean.p.accept.no.sel, score)
r.score.CN = cor(score, CN)

cor.df = data.frame("comparison" = c("p.accept.no.sel vs CN", "p.accept.no.sel vs score", "score vs CN"),
                    "R" = c(r.mean.p.accept.no.sel.CN , r.mean.p.accept.no.sel.score, r.score.CN))

knitr::kable(cor.df, digits = 2, caption = "Correlation between sequence profiles")

par(pty = "s")

plot(x = log(mean.p.accept.strong.sel), y = CN)
plot(x = log(mean.p.accept.strong.sel), y = score)
plot(x = score, y = CN)

r.mean.p.accept.strong.sel.CN = cor(log(mean.p.accept.strong.sel), CN)
r.mean.p.accept.strong.sel.score = cor(log(mean.p.accept.strong.sel), score)
r.score.CN = cor(score, CN)

cor.df = data.frame("comparison" = c("log(p.accept.strong.sel) vs CN", "log(p.accept.strong.sel) vs score", "score vs CN"),
                    "R" = c(r.mean.p.accept.strong.sel.CN , r.mean.p.accept.strong.sel.score, r.score.CN))

knitr::kable(cor.df, digits = 2, caption = "Correlation between sequence profiles")

```
